\pisodm: Building SOC Applications in the Presence of Non-Functional Requirements

### 2017-10-14 JournalPISODM2016/SubmittedV2/AnswersToReferees.tex

1.	On Section 2, the related works are cited as such. This section lucks of an evaluation of these works, and of the motivation of your proposed method. 
\end{quotation}

\noindent 
We added 2 paragraphs at the end of Section 2: (i) considering the evaluation of the related work; and (ii) also motivating our proposed method. 
\begin{quotation}\sf\footnotesize

2.	This method can be fully automated? Is it appropriate to non-IT professionals? 
\end{quotation}

\noindent 
We highlighted at the beginning of section 4 (page 15), that the method is semi automatic. We explain the reasons the method is not completely automatic.

@@

\noindent 
We highlighted in section 3.2.2, how the service is chosen.  To be chosen and executed the service must fit the contract (constraints). If the service fits the contract it will be used in the execution flow.

\begin{quotation}\sf\footnotesize

4.	The paper validates the method with an industrial use case, but it lacks of a real evaluation that shows how the proposed method improves the existing once. For instance, evaluating the satisfaction of the users of the method, and comparing their satisfaction with the use of another development method. 
\end{quotation}

\noindent 
\hl{We intend to evaluate our solution as suggested by the reviewer in our future work. That said, it is worth mentioning that the effectiveness of the solution proposed was showcased in the paper using a real-world application that underlines the features of our solution. In page 24 Section 5.3  we discussed about the importance of the use case for validating the methodology and about the perspectives of evaluating (i) time to market when a new module will be designed and implemented in the system (ii) maintainability of the new module having a separated and clear vision of business logic and non-functional properties.}

@@

\begin{quotation}\sf\footnotesize

5.	How the integration and the description of the constraints (Non-functional properties) improves the development process and the developed application? 
\end{quotation}

\noindent 
\hl{ We believe that across the paper we have shown that decoupling  the description (specification) of non functional properties from the functional aspects, provides developers with the means to focus on the logic of their application. This isolates them from the nitty-gritty aspects of non-functional aspects. Furthermore, it improves modularity by allowing the same application to be run under different non-functional settings, if needed. We explained this on page 24 paragraph starting with "We learned ..."
}

@@

\begin{quotation}\sf\footnotesize

6.	These non-functional requirements can include other types of constraints? 
\end{quotation}

\noindent 
As the non-functional requirements concept used in our method is related to the following types of constraints (page 9): (i) business rules (constraints / contract rules) associated to the behaviour of the application (services); (ii) Value Constraint related to the data types associated to attributes (like in the relational model); (iii) exceptional behaviour associated to the behaviour of the application when an exception is produced.

%it is possible include different types of constraints. The constraints are defined as pre- and post-conditions relates to services. So, during the design process, it is possible to include different kind of contracts related to services.

Figure 4 ($\pi$-UseCase Meta-Model.) presents the Policy View Concepts used to design the constraints.

@@

\begin{quotation}\sf\footnotesize

7.	On page 13, you specify that the model can be automatically translated into actual computer programs. Is this actually developed? If not, how is that can be possible? 
\end{quotation}

\noindent 
%Yes, it is automatically translated into $\pi$-PEWS \cite{MPC08,BaCAM05,BHM06,BHM06rep} specification programs. It is possible because of the PIM to PSM transformation rules described in sections 4.3 and 4.4 (using Acceleo specification programs) and also described in \cite{placidoPhDThesis2012}.

 Yes, at this time the $\pi$-PEWS model can be automatically translated into a $\pi$-PEWS executable program (this is already implemented using Acceleo platform).

@@

\noindent 
Fixed in the new version.

\begin{quotation}\sf\footnotesize

+ The method is an extension of SOD-M, but I was wondering why the new name is piSOD-M. If an explanation exists, it would be interesting to indicate (since there is nothing to do with pi-calculs sometimes used in the WS world) 
\end{quotation}

\noindent 
The motivation behind the name has nothing to do with $\pi$ - calculus which is indeed used for formalizing service coordinations. $\pi$ comes from Policy which is the central notion to represent  NFR's and the way they are enforced at runtime. As you have seen, NFR are at some point implemented by sets of Event - Reaction rules. In the domain of active databases sets of rules could define a Policy. $\pi$ stands of the pronunciation of the first letter of the word "policy" which is /p/. The explanation of the name appears in a footnote of page 1 as explanation, the first time the name $\pi$-SOD-M appears in the text.

%It just sounded appropriate for an extension of the existing SOD-M approach.

### 2017-10-14 JournalPISODM2016/SubmittedV2/main.tex
Valeria's comments

\begin{abstract}
Specifying non-functional requirements (NFRs) is a complex task, being usually addressed during the latter phases of the software development process.
The late inclusion of NFRs during the software development  may compromise the quality of the resulting application.
This paper presents \pisodm, a method and associated tools that
\textit{(i)}  allow the early specification of NFRs allowing users to them in an abstract way without having to  care about low level details;
\textit{(ii)} embraces the MDA philosophy, generating models (code) whenever possible for reducing the specification and programming effort of the business application logic and its associated NFRs.
Our solution has been used in the context of an industrial and real case study.
\end{abstract}

@@

Figure~\ref{fig:CIM:usecasemetamodel} shows the \textit{$\pi$-UseCase} meta-model.
This model is  to be the first representation of the application in terms of functionality, as well as to represent its non-func\-tion\-al requirements.
The notion of \textit{policy} is used to describe NFRs.
Figure~\ref{fig:CIM:usecasemetamodel} highlights the concepts related to NFRs.
The \textit{$\pi$-UseCase} meta-model extends the UML Use Case meta-model to describe NFRs with  the following  concepts:  {\sc Business service}\footnotemark \footnotetext{\hl{We use the 
%{\sc 
small caps
%} 
font for referring to classes of a meta-model.}}, {\sc End consumer}, {\sc Requirement}, {\sc Use Case}, {\sc Composite Use Case}, {\sc Non-\-Func\-tion\-al Requirement}, {\sc Non-Functional} {\sc At\-tri\-bute} and {\sc Constraint}.
An {\sc End Consumer} is represented by an {\sc Actor}.
An  {\sc Actor} is related to {\sc Use Cases} (from the original UML definition), as a {\sc  Composite Use Case} is a set of actions performed by the system which can be broken into different {\sc Use Cases}.
{\sc  Business Service} aggregates several {\sc Use Cases},  a service can be expressed by one or more use cases.
 The {\sc Business Collaborator} concept is represented  by {\sc Packages}.
A {\sc Business Collaborator} denotes an external service or a system that interacts with the application that is being modeled.
Each {\sc Business Collaborator} combines the features described in each  {\sc Package}.
%Thus, the service functions can be specified as being grouped into packages.

@@

 The {\sc Non-functional requirement} and {\sc Non-functional Attribute} concepts are represented by {\sc Use Cases} and {\sc Constraints}.
A {\sc Use Case} may have several {\sc Constraints}. Each {\sc Constraint} has a name, description, and a flag to indicate that it has to be  dynamically checked.
Each {\sc Con\-straint} is represented as a stereotyped ({\sf constraint})\footnotemark \footnotetext{\hl{We use the 
%{\sf 
sans serif
%}
 font for referring to classes defined using a meta-model.}} use case.
There are three kinds of constraints:
\textit{(i)} A data pair ({\sc Value Constraint}), represented as the stereotype ``value'';
\textit{(ii)} Business rules (\textsc{Business Constraint}), represented as the stereotype {\sf ``business''}; and
\textit{(iii)} An {\sc Exceptional Behavior} constraint, represented by the stereotype {\sf ``exceptional\_behavior''}.
The use of these constraints is shown in the next example.

@@

\pisodm  proposes intra-level transformation rules from $\pi$-use case to $\pi$-service process models, $\pi$-service process to $\pi$-service composition  models, as well as inter-level transformation rules from $\pi$-service composition to $\pi$-PEWS models. 
\hl{These rules were implemented into a semi-automatic tool} (the \pisodm  plugin\footnotemark \footnotetext{\hl{At each stage of transformation it is possible to include new constraints. In the more abstract stages of transformation there is complete automation in the models (i.e., from $\pi$-UseCase to $\pi$-ServiceProcess, and from $\pi$-ServiceProcess to $\pi$-ServiceComposition). On the other hand, in the case of the transformation to specific platforms (i.e. from $\pi$-ServiceComposition to $\pi$-PEWS), a review by the analyst is required after the generation of the specification code. Thus, the method is more appropriate for those who have background knowledge in Information Technology.}} for Eclipse).
The transformations from CIM to PIM level models are not automatized, due to the informal nature of the CIM level.

@@

Finally, the system FlyingPig had been incrementally built using classic software engineering techniques before applying $\pi$SOD-M. It is difficult to compare the initial time to market versus the time to market using our method. The company could not estimate a realistic value of this measure for the first version of the application. We intend to evaluate the quality of experience of programmers, when they will have to build a new version of FlyingPig. The quality of experience related to $\pi$SOD-M will be given by measuring time to market and maintainability which are the two measures important to the company. }

>>>>

\begin{abstract}
Specifying non-functional requirements (NFRs) is a complex task, being usually addressed during the latter phases of the software development process.
The late inclusion of NFRs during the software development  may compromise the quality of the resulting application.
This paper presents \pisodm \footnote{\hl{The letter $\pi$  of the name of the method comes from Policy which is the central notion to represent  NFR's and the way they are enforced at runtime.  $\pi$ stands of the pronunciation of the first letter of the word "policy" which is /p/.}}, a method and associated tools that
\textit{(i)}  allow the early specification of NFRs allowing users to them in an abstract way without having to  care about low level details;
\textit{(ii)} embraces the MDA philosophy, generating models (code) whenever possible for reducing the specification and programming effort of the business application logic and its associated NFRs.
Our solution has been used in the context of an industrial and real case study.
\end{abstract}

@@

Figure~\ref{fig:CIM:usecasemetamodel} shows the \textit{$\pi$-UseCase} meta-model.
This model is  used to give the first representation of the application in terms of functionality, and of its non-func\-tion\-al requirements.
The notion of \textit{policy} is used to describe NFRs.
Figure~\ref{fig:CIM:usecasemetamodel} highlights the concepts related to NFRs.
The \textit{$\pi$-UseCase} meta-model extends the UML Use Case meta-model to describe NFRs with  the following  concepts:  {\sc Business service}\footnotemark \footnotetext{\hl{We use the 
%{\sc 
small caps
%} 
font for referring to classes of a meta-model.}}, {\sc End consumer}, {\sc Requirement}, {\sc Use Case}, {\sc Composite Use Case}, {\sc Non-\-Func\-tion\-al Requirement}, {\sc Non-Functional} {\sc At\-tri\-bute} and {\sc Constraint}.
An {\sc End Consumer} is represented by an {\sc Actor}.
An  {\sc Actor} is related to {\sc Use Cases} (from the original UML definition), as a {\sc  Composite Use Case} is a set of actions performed by the system which can be broken into different {\sc Use Cases}.
{\sc  Business Service} aggregates several {\sc Use Cases},  a service can be expressed by one or more use cases.
 The {\sc Business Collaborator} concept is represented  by {\sc Packages}.
A {\sc Business Collaborator} denotes an external service or a system that interacts with the application that is being modelled.
Each {\sc Business Collaborator} combines the features described in each  {\sc Package}.
%Thus, the service functions can be specified as being grouped into packages.

@@

 The {\sc Non-functional requirement} and {\sc Non-functional Attribute} concepts are represented by {\sc Use Cases} and {\sc Constraints}.
A {\sc Use Case} may have several {\sc Constraints}. Each {\sc Constraint} has a name, description, and a flag to indicate that it has to be  dynamically checked.
Each {\sc Con\-straint} is represented as a stereotyped ({\sf constraint})\footnotemark \footnotetext{\hl{We use the 
%{\sf 
sans serif
%}
 font for referring to classes defined using a meta-model.}} use case.

\hl{There are three kinds of constraints:}
\textit{(i)} A data pair ({\sc Value Constraint}), represented as the stereotype ``value'';
\textit{(ii)} Business rules (\textsc{Business Constraint}), represented as the stereotype {\sf ``business''}; and
\textit{(iii)} An {\sc Exceptional Behaviour} constraint, represented by the stereotype {\sf ``exceptional\_behavior''}.
The use of these constraints is shown in the next example.

@@

\pisodm  proposes intra-level transformation rules from $\pi$-use case to $\pi$-service process models, $\pi$-service process to $\pi$-service composition  models, as well as inter-level transformation rules from $\pi$-service composition to $\pi$-PEWS models.

\hl{These rules were implemented into a semi-automatic tool, the $\pi$-SOD-M  plug in for Eclipse. At each stage of transformation it is possible to include new constraints.} 
\hl{Transformation in the more abstract stages is completely automatic  (i.e., from $\pi$-UseCase to $\pi$-ServiceProcess, and from $\pi$-ServiceProcess to $\pi$-ServiceComposition). In contrast, the transformation to specific platforms (i.e. from $\pi$-ServiceComposition to $\pi$-PEWS), the analyst must intervene and check the generation of the specification code because there are programming details that can be interpreted in different manners depending on the target platform. In this sense, the method is more appropriate for those who have background knowledge in Information Technology.}.
The transformations from CIM to PIM level models are not automatized, due to the informal nature of the CIM level.

@@

Finally, the system FlyingPig had been incrementally built using classic software engineering techniques before applying $\pi$SOD-M. It is difficult to compare the initial time to market versus the time to market using our method. The company could not estimate a realistic value of this measure for the first version of the application. We intend to evaluate the quality of experience of programmers, when they will have to build a new version of FlyingPig. The quality of experience related to $\pi$SOD-M will be given by measuring (i) time to market when a new module will be designed and implemented in the system (ii) maintainability of the new module having a separated and clear vision of business logic and non-functional properties. Both two measures  are important to the company. }
